\hypersetup{
  linkcolor=black,
  linkbordercolor={0 0 0}
}

\def\LayoutTextField#1#2{\makebox[6em][l]{#1}\raisebox{-.5ex}{#2}}
\def\LayoutChoiceField#1#2{\makebox[6em][l]{#1}#2}
\newdimen\longline
\longline=\textwidth\advance\longline-6em

\makeatletter
\patchcmd{\HyField@FlagsRadioButton}{\HyField@SetFlag{Ff}{Radio}}{}{}{}
\makeatother
\def\DefaultOptionsofRadio{print}

\newcommand{\quest}[3]{
  \medskip
  \begin{minipage}{\linewidth}
    \noindent
    #2
    \medskip
    \begin{adjustwidth}{-2cm}{-2cm}
      \begin{center}
        \ChoiceMenu[radio,name=#1]{}{#3}\\
      \end{center}
      \bigskip
    \end{adjustwidth}
    \end{minipage}
  \bigskip
}

\begin{center}
    \bf Docula: Evaluation Questionnaire
\end{center}
The rest of this document contains some questions that will help me to
evaluate the functionality of \verb|docula|, please take the time to fill them
out; all non-multiple choice questions are optional.

\begin{Form}
  \quest{current}{
    Do you currently use any documentation generation software? If so, please
    state what you use.
  }{Yes, No}
  \medskip
  \TextField[name=additionalcomment,width=\longline,
    borderwidth=0,backgroundcolor={.85 .85 .85}]{}\\

  \quest{current_easy}{
    How do you find the current software to use?
  }{Very Difficult, Difficult, Fair, Easy, Very Easy}

  \quest{current_output}{
    How would you rate the output of the software that you currently use?
  }{(Lowest) 1, 2 , 3 , 4 , 5 , 6 , 7 , 8 , 9 , 10}

  \quest{expectations}{
    What are your expectations of the software you are about to test?
  }{(Lowest) 1 , 2 , 3 , 4 , 5 , 6 , 7 , 8 , 9 , 10}

  Along with this questionnaire you should have received a .gem file; to allow
  for convenient use of docula whilst testing it functionality, please execute
  the following in a shell\footnote{Depending on your Ruby setup, this may need
  to be executed as root}:
  \begin{verbatim}
  $ gem install --local docula-*.gem
  \end{verbatim}
  This will properly install docula and its dependencies onto your system. You
  should then be able to use the \verb|docula| command to generate documentation
  for a particular directory. The resulting documentation should be available in
  the \verb|output/| directory by default.

  In addition to the .gem file, you should also have received a .tar.gz file
  containing some C source files; please extract this archive and navigate to the
  created directory in a terminal, then run the command:
  \begin{verbatim}
  $ docula
  \end{verbatim}
  \verb|docula| should now parse the files in the directory, you may see some
  status messages as it executes; when \verb|docula| has finished running, you
  will be able to view the generated documentation by opening
  \verb|output/files.html| in a web browser.\\

  \quest{success}{
    Did docula successfully produce documentation?
  }{Yes, No}

  \quest{speed}{
    How would you rate the speed with which docula completed its
    execution?
  }{Very Slow, Slow, Fine, Fast, Very Fast}

  Please review the generated documentation; \verb|files.html| will contain a list
  of the files parsed, and clicking on one of the files will bring you to the
  documentation for the contents of that file.

  Try to locate the function used to fetch the description of an audio channel,
  using the documentation provided.\\

  \quest{locate}{
    Are you able to determine what the function returns in the event of an
    error?
  }{Yes, No}

  \quest{locate_speed}{
    How quickly were you able to find the function?
  }{Very Slowly, Slowly, As Expected, Quickly, Very Quickly}

  \quest{navigate}{
    In general, how did you find navigating the documentation?
  }{Very Difficult, Difficult, Fine, Easy, Very Easy}

  The file \verb|time.h| is lacking documentation, please add some documentation
  to the two functions and then re-run \verb|docula|.\\

  \quest{changes}{
    Were your changes successfully detected \& displayed?
  }{Yes, No}

  \quest{appear}{
    Did they appear as you would have expected?
  }{Yes, No}

  \quest{compare_speed}{
    How did the speed of execution compare with the last time you ran docula?
  }{Much Slower, Slower, About The Same, Quicker, Much Quicker}

  \quest{appearance}{
    How would you rate the appearance \& layout of the documentation?
  }{Terrible, Poor, Average, Good, Excellent}

  \quest{compare_expect}{
    How did the software compare to your expectations?
  }{Much Worse, Worse, About Right, Better, Much Better}

  \noindent
  Was there anything in particular that you felt was missing from the
  documentation?\\
  \medskip
  \TextField[name=missingcomment,multiline=true,width=\longline,height=1.75in,
    borderwidth=0,backgroundcolor={.85 .85 .85}]{}\\

  \bigskip
  \noindent
  Additional Comments\\
  \medskip
  \TextField[name=additionalcomment,multiline=true,width=\longline,height=2in,
    borderwidth=0,backgroundcolor={.85 .85 .85}]{}\\
\end{Form}

\begin{center}
  \bigskip\hrule\bigskip
  \emph{Thank you for taking the time to complete this questionnaire.}
\end{center}
