% Principally, this chapter should describe the work which was undertaken before
% code was written, hardware built or theories worked on. It should show how the
% project proposal was further refined and clarified, so that the Implementation
% stage could go smoothly rather than by trial and error.

% Throughout this chapter and indeed the whole dissertation, it is essential to
% demonstrate that a proper professional approach was employed.

% The nature of this chapter will vary greatly from one dissertation to another
% but, underlining the professional approach, this chapter will very likely
% include a section headed "Requirements Analysis" and incorporate other
% references to the techniques of Software Engineering.

% The chapter will cite any new programming languages and systems which had to
% be learnt and will mention complicated theories or algorithms which required
% understanding.

\chapter{Preparation}
\section{Requirements Analysis}

The requirement of this project is to produce a program that can successfully
generate documentation for source code and that groups the resulting
documentation in a logical manner.

This can be broken down into two smaller targets:
\begin{enumerate}
  \item A section capable of parsing code to extract the documentation and
    related information.
  \item A section that transforms the extracted data into a document.
\end{enumerate}

The documentation produced should provide details for all functions documented,
correctly matching up any information provided about arguments with the actual
arguments and their type from the function itself.

Additionally, the program should require no intervention from the user to
complete its generation process.

I also need to determine which language is the appropriate choice for
implementing this project, since each will have their own strengths \&
weaknesses, and how I will go about testing that everything is working as it
should along the way. While deciding on testing strategies before programming is
not required, it is important to test at some point in development (whether it
be throughout, at the end or both) and deciding on it beforehand ensures that it
will not be neglected.

\section{Software Architecture}
This project consists of five parts:
\begin{itemize}
  \item One or more parsers, to transform the code base into an appropriate data
    structure.
  \item Code to process the above data structure and extract the required
    information from the code.
  \item Storage of the intermediate data.
  \item Code to transform the intermediate data into one or more human-readable
    formats.
  \item A command line interface, for use by the user.
\end{itemize}

For the purposes of this project, I will only be implementing a parser for one
programming language; it will need to be a simple, standardised language to be
completed in the time available, there should also be sufficient code available
written in the language to provide thorough test cases. I have chosen to
implement the parser for C, as it is relatively easy to parse, and there is
plenty of example code available on the internet.

Likewise, I will be implementing one concrete output format for the program; the
format should be one I am already familiar with and should easily support
linking between sections. I have decided that HTML is the best fit for the
output format, since it is easy to create \& does not require any extra tools to
view the final result.

Even though I am only implementing one of each of the above stages, this does
not mean that I will be designing this software without other input \& output
formats in mind; the software should be developed with a view to be expanded in
the future, and therefore should be coded to be as modular as possible.

Reprocessing all the code in a large project can become time-consuming; the data
storage component allows us to generate documentation incrementally, avoiding
large time penalties and allowing for documentation to be available in
interactive time frames. This can be done by determining whether a file has
changed since it was last parsed, if it has not there is no need to reinterpret
the file, we can simply use the data that is already stored about it. This is
particularly useful, as it will allow for quick incremental generation of the
documentation, so developers using the program will find it much more convenient
to quickly use to check that their documentation is in order.

I will be following the waterfall model of development for this project, as far
as is sensible, as maintenance is outside the scope of this project.

\begin{center}
  \vspace*{5mm}
  \begin{tikzpicture}[node distance = 2cm, auto]
    % Place nodes
    \node [block_spaced] (input) {input};
    \node [block_spaced, right=of input] (processing) {data processing};
    \node [block_spaced, above=of processing] (storage) {data storage};
    \node [block_spaced, right=of processing] (output) {output formatting};
    \node [block_spaced, below=of processing] (command) {application logic};
    % Draw edges
    % \path [line] (input) -- (processing);
    \path [line, dashed] (processing) -- (storage);
    \path [line, dashed] (storage) -| (output);
    \path [line] (command) -- (input);
    \path [line] (command) -- (processing);
    \path [line] (command) -- (output);
  \end{tikzpicture}
\end{center}

The application component will act not only as an interface with the user, but
also interact with each of the input, processing \& output components, calling
them with the appropriate options and so on. This reduces coupling between the
other stages, since the command line component is the only one that needs
knowledge of all the stages, allowing for increased modularity of the code.

\section{Comment Format}
In order to produce documentation from source code I need to define what
constitutes documentation, both so that I can parse \& process it and so that
appropriate documentation is actually written. This means defining a comment
format, I have previously used both Javadoc \& doxygen and have reviewed them
again prior to beginning this project. I have therefore kept their functionality
and limitations in mind while designing the comment format.

In my proposal, I specifically mentioned the compatibility with existing comment
formats was highly important to produce a useful result; as such, consideration
about the format I was going to be using had to be undertaken before any work
could begin. Both doxygen and Javadoc's comment formats are similar to one
another, and together they have quite a large market share, so it was sensible
to design a format that was compatible with these as much as possible. Ideally,
my format should be equal to or a superset of their formats, so that any
comments written with them in mind will work with the software I write.

With that in mind, I designed a format along the following lines (a full version
of which can be found in the appendix):

\begin{lstlisting}[language=c, escapechar=~]
  /**
   * Summary
   *
   * Full Description
   *
   * ~{\color{solarized@cyan} @param[in, out]}~ ptr Description of parameter ptr
   * ~{\color{solarized@cyan} @param[in]}~ size Description of parameter size
   * ~{\color{solarized@cyan} @return}~ Description of return value.
   */
  void *realloc(void *ptr, unsigned int size);
\end{lstlisting}

\begin{itemize}
  \item Asterisks at the beginning of lines within the comment are ignored.
  \item The first paragraph (i.e.~up to the first empty line) is considered to
    be a summary, and gives a brief outline of the intended use. This is
    standard practice with Javadoc, and can be enabled in doxygen.
  \item The following paragraphs make up the full description, expanding on the
    summary, describing in more detail what it does and how it should be used.
  \item @-annotations
    \begin{itemize}
      \item \lstinline|@param| annotations are used to describe the individual
        parameters and their purpose.
      \item The square-bracket suffixes to \lstinline|@param| annotations are
        for describing the flow of data, that is whether the variable is only
        used for input, output or both. These are described as \lstinline|[in]|,
        \lstinline|[out]| \& \lstinline|[in,out]| respectively. If a parameter
        is absent a flow annotation, it will be assumed that it is
        \lstinline|[in]|.
        doxygen already uses this way of describing flows; Javadoc does not have
        any such feature.
      \item \lstinline|@return| is used to describe what will be returned by the
        function; this may include under what conditions it will return an error
        and what those error values may be.
    \end{itemize}
    \item Summary \& Description sections can be applied to things other than
      functions, the @-annotations cannot, as they would be meaningless.
\end{itemize}

\section{Tools}
  \subsection{Programming Language}
    I opted to use Ruby to implement this project, whilst I do not have as much
    experience with Ruby as some other languages (like Java), I felt it was best
    suited to the project. Since one of Ruby's strengths is in its powerful
    text-manipulation capabilities it is well suited to this project.

    Additionally, having looked into implementing parsers in various languages,
    I found that Treetop\cite{website:treetop} was well-suited to my needs.
    Treetop is a domain-specific language for Ruby, that facilitates the
    creation of parsing expression grammars; one of the particularly useful
    things about Treetop is that it allows you to define specific types for
    matched rules to instantiate, so that a lot of the effort of having to walk
    a parse tree can be alleviated. This will be discussed in further detail in
    the Implementation chapter.

    I was already partially familiar with Ruby, however I made use of
    \emph{Eloquent Ruby}\cite{book:eloquent_ruby} to further hone my knowledge
    of the language.
  \subsection{Version Control}
    Given that I have more previous experience with git than with any other
    version control system, I have decided to use it to manage my project. It
    provides some advantages over some other version control systems,
    particularly in the fact that it is distributed; which means that the entire
    development history of the project will be available to me, no matter what
    machine I am working on.

    Using git will also allow me to make use of both GitHub and the PWF as a
    means of backup by pushing my changes to them regularly. As an additional
    form of backup my main development machine is backed up using OS X's Time
    Machine.

\section{Testing}
The testing of the project will be split into parts; part of it, particularly
in the development of the parser, will be performed during development and the
rest will be performed after the completion of the main body of the code has
been written.

  \subsection{Regression Testing}
    In order to implement the parser successfully, I will be making use of
    regression testing; this will be used to ensure that the parser behaves
    in an expected manner for specific inputs, and make sure that when new
    functionality is added it does not fail any of the previous tests.

    To do this I will be writing short snippets of code that implement a
    specific part of the language's feature set, and then making sure that these
    tests pass before continuing, such as:

    \begin{lstlisting}[language=c, gobble=4]
      int c;
    \end{lstlisting}

    This makes sure that I identify bugs early, and prevents me from introducing
    any new ones later in development, allowing me to fix them before they cause
    any serious problems.

  \subsection{User Testing}
    Evaluating documentation is inherently a subjective process, and so it would
    not be feasible to evaluate by any formal means; as such, I will be creating
    a questionnaire which will be given to several members of staff at
    Dr.~Watts' company, along with the software itself, for them to evaluate
    compared to their usual documentation solutions. This questionnaire, and any
    familiarisation with the software required, should take no longer than half
    an hour.

    In order for the surveys to be completed the documentation software will
    have to be used on a sample of code from the workplace ensuring that it
    functions across a wide spectrum of styles and is therefore suitable for use
    in the working environment.

  \subsection{Evaluation}
    In order to properly evaluate the success of the project I will be
    evaluating my program by the following criteria:
    \begin{description}
      \item[Correctness -] Does the program produce correct documentation for
        a test set of projects? This will be evaluated manually, by myself, to
        check that the output produced contains all the information expected.
      \item[Time -] Does the program execute quickly enough to be a viable
        documentation solution?
      \item[Functionality -] Is the program actually useful to users? This will
        be evaluated using the results collected in the user testing.
    \end{description}


% \section{Action Plan}
% I then created an action plan with these main aims in mind in order to ensure
% the project's success; I will be be adopting the waterfall model to develop this
% project, and so the plan follows that general structure.
% \begin{enumerate}
%   \item Research - look at existing documentation software, and identify their
%     flaws \& attributes.
%   \item Identify the features required in this parser, taking into account the
%     strengths \& weaknesses identified in the previous phase to determine the
%     functionality to be included or improved upon.
%   \item Develop an outline for the project so that an awareness of structure and
%     flow can form. Plan what order the code shall be written in and what order -
%     what tests need to be performed along the way to ensure functionality.
%   \item Choose the programming language best suited to the project - learn
%     and/or practise the language to ensure fluency and suitability for the
%     project.
%   \item List and develop tests that will take place during the coding and at the
%     end to ensure the project functions correctly.
%   \item Begin writing and testing code - revisit plan if necessary to make
%     alterations.
%   \item Complete end testing and make any necessary alterations
% \end{enumerate}
